RegressionPvNP

### 2017-04-21 RegressionPvNP.Rmd
Minor bugs

```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
pi.hat <- function(x) exp(x)/(1+exp(x))

precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
  }

@@

  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
    message(paste(i, "fold"))
    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
    test = x[test.rows,]
    training = x[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1)

    pred.class <- predict(fit, test, type='class') # check this later
    auc[i,] = apply(pred.class, 2, function(pd) {
      if (sum(as.numeric(pd)) <= 0) return(0)
      as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
    })

    # careful: matrix
    pred <- pi.hat(predict(fit, test)) 
    cv.class[i,] = apply(pred, 2, function(p) {
      # Confusion matrix
      #cmat = table(labels[test.rows], p>0.5)
      #sum(diag(cmat))/sum(cmat)
      (1-p)%*%labels[test.rows] # quantitiative
    })
  }
  return(list('cv.class'=cv.class, 'auc'=auc))
}

@@

cv.patient = cv.patient.glmnet(dysplasia.df, labels, lambda, patients)
lambda.opt <- lambda[which.min(colMeans(cv.patient$cv.class))]

df = cbind.data.frame('lambda.at'=1:ncol(cv.patient$cv.class),'mean'=colMeans(cv.patient$cv.class), 
                      'sme'=apply(cv.patient$cv.class, 2, sd)/5, 
                      'sd'=apply(cv.patient$cv.class, 2, sd), 
                      'lambda'=lambda)
smeA = sd(colMeans(cv.patient$cv.class))/ncol(cv.patient$cv.class) #??

ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
  geom_point(data=subset(df, mean == min(mean)), aes(y=mean, x=lambda.at), col='red') +
  geom_point(data=df[which.min(df$mean)+1,], aes(y=mean, x=lambda.at), col='blue') +
  geom_point(data=df[which.min(df$mean)-1,], aes(y=mean, x=lambda.at), col='green')

lambda.opt = lambda[which.min(df$mean)-1]

coef.opt <- coef.glmnet(fit0, s=lambda.opt)
coef.opt = as.matrix(coef.opt)
coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])

@@

At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.  Using these to fit a linear model by removing the remainder of the features we get a pretty clear separation between samples from progressors and non.

```{r echo=T, tidy=T, warning=F, message=F}

# So then the best model uses a lambda value in this range?
bestF = rownames(coef.opt)

## So the model would be...
data = dysplasia.df[,bestF]

fit = lm(labels~., data=cbind.data.frame(data, labels))
summary(fit)
```

```{r echo=F, warning=F, message=F}
df = cbind.data.frame('index'=1:length(labels), 'fitted'=fit$fitted.values, 'status'=labels, 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels == 1, fit$fitted.values > 0.5), 3),
            'Non-Progressors'=signif(precisionRecall(labels == 0, fit$fitted.values < 0.5),3))
```

The model separates surprisingly well (assuming this was the correct setup).

`r pander( pc, caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

## Questions

While this looks great, I'm a bit suspicious that it's too good.

1. Can it still predict well if all HGD/IMC samples are removed from the matrix?
2. If I remove one sample from the progressors and iteratively rerun it, at what point does it fail?  I would expect this to occur at least at the earliest timepoints where BE is indicated.

@@

Just as a point, the initial model is tested against `r nrow(data)` samples with `r ncol(data)` features.

@@

samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)

fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
```

This model is fit against `r nrow(data[samples,])` samples instead as `r nrow(data)-length(samples)` are pathologically HGD/IMC. While this looks great, I'm a bit surprised.

```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}

df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
```

`r pander(pc , caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

### Remove LGD Samples as well

This just strikes me as far too good.  I would expect some of the progressor samples to separate even with LGD removed, but not all.

```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}

samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

`r pander(pc,caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

### Peel back samples one at a time until it fails

This is not what I would expect. By the time I get to the earliest timepoint most of the remaining samples in the progressors are BE only. Some of them should not separate well.

```{r peel, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
get.next.samples<-function(patient.data, i) {
  as.vector(unlist(sapply(patient.data, function(pd) {
    j = i
    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)

@@

nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
for (i in seq(1,35,5)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
  samples = c(nonps, ps)

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
  summary(fit)

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
    #ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    #  labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

  pander(pc, justify='left')
}

@@

`r pander(table(subset(patient.info, Samplename %in% samples)[,c('Status','Pathology')]), caption="Pathology of samples in final model")`

## Again randomly assigning labels

One check is to randomly assign labels and rerun the model, at the same time I am pulling out samples in time backwards. The good thing is that the model no longer works as I would expect.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
fstats = matrix()
nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
for (i in seq(1,35,10)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
  samples = c(nonps, ps)

  randLabels = sample(c(0,1), length(samples), replace = T)

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=randLabels))
  summary(fit)

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

  pander(pc, justify='left')
}
```

Final precision recall, is just too good:

`r pander(pc, justify='left')`

# Biases?

Some of the samples are going to have duplicated values in the matrix due to the method of overlapping segments to encode all of the patient data together.  It's possible that is causing the progressors to stand out more?

### 1. Duplication of values - remove samples

If I remove samples (not patients) by the number of duplicated values they have in the matrix the fit actually separates more. Somehow I expected they might separate less.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
duprows = apply(data, 1, function(row) {
  length(which(duplicated(row)))
})

pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])

#t.test(subset(pi, Status == 'NP')$duprows, subset(pi, Status == 'P')$duprows)$p.value

## So what happens if I remove from the model those with the most duplicated rows? 
#summary(duprows)

samples = names(duprows[which(duprows <= median(duprows))])

pander(
  cbind(pi %>% group_by(Status) %>% summarise('total.dup'=sum(duprows,na.rm=T)), table(labels[samples]))[,c(1,2,4)],
  caption="There are significantly more duplicated rows following the overlap process in progressors (total.dup). In the samples that have less than the median number of duplicated values, both P and NP are still represented at about the same rate (~44% of rows are P)")

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

`r  pander(pc, caption=paste('The median for duplicated values in a row is',median(duprows),'the linear model that uses just those rows with less than the median shows a complete separation'))`

### 2. Duplication of values - remove by patient

Remove all samples from single patients if any of their samples have a high number of duplicated rows, most of these are progressors. The resulting fit is still suspiciously good though so I think the bias isn't based on duplicated values.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])

## So what happens if I remove from the model those with the most duplicated rows? 
#summary(duprows)

toremove = names(duprows[which(duprows > median(duprows))])

samples = intersect(rownames(data), subset(patient.info, Patient %in% unique(subset(patient.info, Samplename %in% toremove)$Patient))$Samplename)

pander(table(labels[samples]), caption="0=NP, 1=P")

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

`r pander(pc, caption=paste('In this I removed all samples from patients that had greater than the median value',median(duprows), 'of duplicated rows. The separation is less perfect but still very good.'))`

```{r, echo=F, warning=F, message=F, eval=F}
test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
test = dysplasia.df[test.rows,]
training = dysplasia.df[-test.rows,]

##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case
library(caret)

objControl <- trainControl(method='cv', number=5, returnResamp='none')
#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)
objModel <- train(dysplasia.df[,bestF],as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)

vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
results$VariableName <- rownames(vimp)
colnames(results) <- c('VariableName','Weight')
results <- results[(results$Weight != 0),]

which(results$VariableName %in% bestF)

ggplot(subset(results, Weight > 10), aes(x=reorder(VariableName, Weight), y=Weight)) + 
  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))

## Very close to the same for the top N coefficients
rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)

```{r labels2, echo=F, message=F, include=F, eval=F}
# y = HGD (1) vs BE/LGD (0)

#Do the same again, but with HGD/IMC labeled instead

## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

# sort in label order
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")
```

```{r cvglmnet, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
#Note on the labels. When trying to classify samples that are HGD or not we have `r table(labels)[1]` samples without and only `r table(labels)[2]`) with HGD.

## cv.glmnet?

gp = plot.multi.cv.glmnet(dysplasia.df, labels, alpha=c(0,0.5,1), title='cv.glmnet HGD vs BE')
do.call(grid.arrange, c(gp, ncol=2, nrow=3))

cv1 = cv.glmnet(dysplasia.df, labels, alpha=1, family='binomial')
plot(cv1$glmnet.fit, main='alpha=1 glmnet.fit from cv.glmnet')
```

```{r xvalpt2, message=F, warning=F, echo=F, fig.height=6, fig.width=6, eval=F}

# This just makes sure the sets don't become too unbalanced with regards to the labels.
sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
s = sample(rep(seq(5), length = length(unique(pts$Patient))))
patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")

while ( length(which(table(patients$group)/nrow(patients) < 0.2)) > 1 ) {
  print(".")
  s = sample(rep(seq(5), length = length(unique(pts$Patient))))
  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
  #sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
}

fit0 <- glmnet(dysplasia.df, labels, alpha = 1) # all patients
plot(fit0)
title(main='fit0, all samples')
lambda <- fit0$lambda

cv.class = matrix(nrow=5, ncol=length(lambda))
prec.rec = matrix(nrow=5, ncol=length(lambda))
auc = matrix(nrow=5, ncol=length(lambda))
for (i in 1:5) {
    test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == i)$Samplename)
    test = dysplasia.df[test.rows,]
    training = dysplasia.df[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, labels[-test.rows], lambda=lambda, family='binomial', alpha=1) 
    plot(fit)
    title(main=paste('fold', i))

    # careful: matrix
    pred <- pi.hat(predict(fit, test)) 
    pred.class <- predict(fit, test, type='class') # check this later

    auc[i,] = apply(pred.class, 2, function(x) {
      if (sum(as.numeric(x)) <= 0) return(0)
      as.numeric(auc(roc(response=as.numeric(x), predictor=labels[test.rows])))  
    })

    # prec.rec[i,] = apply(pred.class, 2, function(p) {
    #    precisionRecall(labels[test.rows], as.integer(p))$recall
    # })
    cv.class[i,] = apply(pred, 2, function(p) {
      # Confusion matrix
      #boxplot(p~labels[test.rows], main=i)
      # cmat = table(labels[test.rows], p>0.5)
      # sum(diag(cmat))/sum(cmat)
      (1-p)%*%labels[test.rows] # quantitiative
    })
    #s = sample(1:ncol(pred),1)
    #boxplot(pred[,s]~labels[test.rows], main=s)
}
lambda.opt <- lambda[which.min(colMeans(cv.class))]

## TODO...NOPE, this pulls nearly all of the features so START OVER.  This is likely because there are so few HGD samples

print(paste("mean auc at optimal lambda:", round(mean(auc[,which.min(colMeans(cv.class))]),2)))

#coef.opt <- coef.glmnet(fit0, s=min(fit0$lambda))
coef.opt <- coef.glmnet(fit0, s=lambda.opt)
coef.opt = as.matrix(coef.opt)
coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])

print( rownames(coef.opt) )

test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
test = dysplasia.df[test.rows,]
training = dysplasia.df[-test.rows,]

##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case

objControl <- trainControl(method='cv', number=5, returnResamp='none')
#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)

predictions <- predict(object=objModel, test)
print(roc(predictions, as.integer(labels[test.rows]))$auc)

#objModel <- train(dysplasia.df,as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)

vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
results$VariableName <- rownames(vimp)
colnames(results) <- c('VariableName','Weight')
results <- results[(results$Weight != 0),]

ggplot(subset(results, Weight > 20), aes(x=reorder(VariableName, Weight), y=Weight)) + 
  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))

## Very close to the same for the top N coefficients
rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)

>>>>

```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
  }

pi.hat<-function(x) exp(x)/(1+exp(x))

non.zero.coef<-function(fit, s) {
  cf =  as.matrix(coef(fit, s))
  cf[which(cf != 0),][-1]
}

@@

  fit.e = list()
  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
    message(paste(i, "fold"))
    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
    test = x[test.rows,]
    training = x[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1) 
    # autoplot(fit) + theme(legend.position="none")

    pred.class <- predict(fit, test, type='class') # check this later
    auc[i,] = apply(pred.class, 2, function(pd) {
      if (sum(as.numeric(pd)) <= 0) return(0)
      #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
      precisionRecall(labels[test.rows], as.numeric(pd))$F1
    })

    # careful: matrix
    pred <- pi.hat(predict(fit, test)) 
    cv.class[i,] = apply(pred, 2, function(p) {
      # Confusion matrix
       #cmat = table(labels[test.rows], p>0.5)
       #sum(diag(cmat))/sum(cmat)
      # quantitiative, pos results + neg results / number of test rows
      (p%*%labels[test.rows] + (1-p) %*% (1-labels[test.rows]))/ length(test.rows)
    })

    fit.e[[i]] = fit
  }

  df = cbind.data.frame('lambda.at'=1:ncol(cv.class),'mean'=colMeans(cv.class), 
                        'sme'=apply(cv.class, 2, sd)/5, 
                        'sd'=apply(cv.class, 2, sd), 
                        'lambda'=lambda)

  lambda.opt = lambda[which.max(df$mean)-1]
  nzcf = lapply(fit.e, non.zero.coef, s=lambda.opt)

  gp = ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
          geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
          geom_point(data=subset(df, mean == max(mean)), aes(y=mean, x=lambda.at, colour="max"), size=3 ) +
          geom_point(data=df[which.max(df$mean)+1,], aes(y=mean, x=lambda.at, colour="max+1"), size=3) +
          geom_point(data=df[which.max(df$mean)-1,], aes(y=mean, x=lambda.at, colour="max-1"), size=3) +
          geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
          annotate("text", x=5, y=max(df$mean)+0.0005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
          theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
          labs(title="CV Performance")

  return(list('cv.class'=cv.class, 'f1'=auc[which.max(df$mean)-1], 'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
}

@@

cv.patient = cv.patient.glmnet(x=dysplasia.df, y=labels, lambda, patients)
lambda.opt = cv.patient$lambda.opt

coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))

grid.arrange(cv.patient$plot)

## check how often the feature(s) is selected at that lamda in each split. "stability selection"

for (i in 1:length(cv.patient$non.zero.cf)) {
  df = as.data.frame(cv.patient$non.zero.cf[[i]])
  colnames(df) = i
  coef.opt = merge(coef.opt, df, by='row.names', all.x=T)
  rownames(coef.opt) = coef.opt$Row.names
  coef.opt$Row.names = NULL
  coef.opt[,(i+1)] = as.integer(!is.na(coef.opt[,(i+1)]))
}
coef.opt = coef.opt[order(sapply(rownames(coef.opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]

@@

`r pander(coef.opt, justify='left', caption='Features from the glm fitted with all data. Each numbered column indicates if the feature was identified in that split.')`

At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.

@@

Just as a point, the initial model is tested against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features.

@@

samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)

cv.nohgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#lambda.opt = cv.patient$lambda.opt
grid.arrange(cv.nohgd$plot + ggtitle("CV Performance, No HGD/IMC samples"))
#coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))

samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#lambda.opt = cv.patient$lambda.opt
grid.arrange(cv.nolgd$plot + ggtitle("CV Performance, No HGD/IMC/LGD samples"))

get.next.samples<-function(patient.data, i) {
  as.vector(unlist(sapply(patient.data, function(pd) {
    j = i
    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)

@@

nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
plots = list()
for (i in seq(1,35,5)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  samples = c(nonps, ps)
  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
}

do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))

@@
